---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->
<!-- usethis::use_readme_rmd() og så Ctlr + Shift + K -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```
# ZIBP-PCA: Zero-inflated Bivariate Poisson Principal Component analysis

This vignette describes how to use the R package zibppca to perform Zero-inflated Bivariate Poisson Principal Component analysis for Neuro-Psychiatric Inventory (NPI) items. 

# Installation

You can install the R package zibppca directly from GitHub. All files in the package have help files, accessed by using ?

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
library(devtools)
devtools::install_github("khellton/zibppca")
?zibppca
```

# Example data 
First the data is loaded as an data.frame. The distribution of items can be visually inspected by histograms or clustered box plots. 
```{r, echo = FALSE,warnings=FALSE}
library(foreign) # to import SPSS sav files
data_new <- read.spss('~/NR/PCANPI/Data/New_selection_nursing_home.sav', to.data.frame=TRUE)
data <- data.frame(data_new[(data_new$aar2004_2011==0),8:19])

get.names <- read.csv("~/NR/ModNPI/Data/NPILong.csv",colClasses = c('factor',rep('numeric',15)),dec =',',na.strings = '#NULL!', header = TRUE,sep = ';')
colnames(data) <- colnames(get.names)[4:15]

head(data, 10L)
hist(data$Delusions_NPI1,main='Delusions')
hist(data$Aggression_NPI3,main='Aggression')
par(mar=c(9,2,0,0))
boxplot(data,las=2,cex.axis=0.7)
```

# Example data 
The zibppca package has the following outline
```{r}
#Load the R package
library(zibppca)

#Run the main command 
result <- zibppca(data,.scale = TRUE)

### Loadings of the first 3 components
result$loadings[,1:3]

### Scree plot of eigenvalues
result$eigenvalues 
plot(result$eigenvalues,main = 'Scree plot',xlab='Eigenvalues')

### Plot patient scores 
plot(result$scores[,1],result$scores[,2],main = 'Score plot',xlab='1. component',ylab='2. component')

```
